### 2021-11-24 input/proposal.tex
meow proposal

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links
    ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually
    reach consistent partitioning if each \process delivers its \TODO{best}
    message among \underline{r}eceived messages $R_x$, and forwards
    its \TODO{best} message among \underline{d}elivered messages $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

  \noindent Whenever a \process $s$ becomes a source, it delivers its
  own message $d_s(\alpha_{s})$. Whatever its set of received
  messages, its own partition $P_s$ becomes and remains its closest
  partition $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq
  s: |\Pi_{s s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $w_{sq} < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\smash{\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s}^{|\Pi_{sq'}|})}$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

A naive implementation resembles echos in acoustics:
\underline{d}elete notifications $\delta$ propagate as long as
receivers' current partition is the targeted one, and when delete
notifications reach the bordering \processes of the deleted partition,
it triggers a competition -- or an echo -- that goes backward to fill
the gap left open by removals.  For example, in Figure~\ref{fig:del},
two partitions initially exist: $P_a$ and $P_d$ that respectively
include $\{a\}$, and $\{b, c, d\}$. In Figure~\ref{fig:delA},
\Process~$a$ deletes its partition. It notifies all neighboring
\processes~--~here only \Process~$b$~-- that may belong to its
partition using \NAMEB. Upon receipt, \Process~$b$ discards the delete
notification $\delta_a$ since $\delta_a$ does not target the former's
partition $P_d$. \Process~$b$ sends its own best partition
$\alpha_d^3$ that may be the best for \Process~$a$. Eventually, every
\process belongs to Partition $P_d$. In this scenario, they reach
consistent partitioning.

@@

\begin{definition}[Happens-before $\rightarrow$~\cite{lamport1978time}]
  The transitive, irreflexive, and antisymmetric happens-before
  relationship defines a strict partial order between events. Two
  messages are concurrent if none happens before the other.
\end{definition}

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest message of a \process matters. A message $m$ conveys
  \underline{s}tale control information if the \process that broadcast
  $m$ later broadcast another message $m'$:
  $\mathcal{S}(m) = \exists m': b_x(m) \rightarrow b_x(m')$.
\end{definition}

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending
($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). In Figure~\ref{fig:problemA}, both $a$ and $c$ become
sources, sending their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. It produces an echo but since $\alpha_c^3$ is already transiting
towards $a$, we simplify it. Finally, Figure~\ref{fig:problemD} shows
that eventually, $c$ stays in the deleted partition $P_a$ for not
having received $\delta_a$ that $b$ blocked earlier. As a cascading
effect, not only $c$ wrongfully believes that it belongs to $a$'s
partition when $a$ already deleted it, but it could contribute to its
inconsistency by blocking farther but up-to-date messages from actual
sources. To reach consistent partitioning, \processes need
\begin{inparaenum}[(i)]
\item to purge all stale messages which
\item implies detecting possibly blocking conditions
\item by acknowledging the last partition of neighbors.
\end{inparaenum}

@@

\begin{lemma}[FB $\wedge$ \underline{F}orwarding of \underline{S}tale
  (FS) $\implies$ Detection]
  \TODO{meow!}
\end{lemma}

>>>>

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

  \noindent \TODO{use total order ($<$) and sets $R_p D_p$} Whenever a
  \process $s$ becomes a source, it delivers its own message
  $d_s(\alpha_{s})$. Whatever its set of received messages, its own
  partition $P_s$ becomes and remains its closest partition
  $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq s:
  |\Pi_{s s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $w_{sq} < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\smash{\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s}^{|\Pi_{sq'}|})}$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

A naive implementation resembles echos in acoustics:
\underline{d}elete notifications $\delta$ propagate as long as
receivers' current partition is the targeted one, and when delete
notifications reach the bordering \processes of the deleted partition,
it triggers a competition -- or an echo -- that goes backward to fill
the gap left open by removals.  For example, in Figure~\ref{fig:del},
two partitions initially exist: $P_a$ and $P_d$ that respectively
include $\{a\}$, and $\{b, c, d\}$. In Figure~\ref{fig:delA}, $a$
deletes its partition. It notifies all neighboring \processes~--~here
only $b$~-- that may belong to its partition using \NAMEB. Upon
receipt, $b$ discards the delete notification $\delta_a$ since
$\delta_a$ does not target the former's partition $P_d$. \Process~$b$
sends its own best partition $\alpha_d^3$ that may be the best for
$a$. Eventually, every \process belongs to Partition $P_d$. In this
scenario, they reach consistent partitioning.

@@

\begin{definition}[Happens-before ($\rightarrow$)~\cite{lamport1978time}]
  The transitive, irreflexive, and antisymmetric happens-before
  relationship defines a strict partial order between events. Two
  messages are concurrent if none happens before the other.
\end{definition}

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only sends messages that it assumes up-to-date:
  $s_x(\alpha_s) \implies \delta_s \not\in R_x$.
\end{definition}

\begin{definition}[\label{def:fd}\underline{F}orwarding of
    \underline{s}taleness (FS)]
  %
  To \NAMEB staleness, a \process $x$ delivers and forwards a received
  staleness notification if it is not yet delivered and targets its
  best delivered message: $\delta_s \in R_x \setminus D_x \wedge
  \alpha_s = \min D_x \TODO{\implies} d_x(\delta_s) \wedge
  f_x(\delta_s)$.
\end{definition}

\begin{definition}[\label{def:echo}\underline{E}chos (E)]
  A \process $x$ that receives a stale message sends back --~or
  echoes~-- its best up-to-date delivered message: $r_{xy}(\alpha_s)
  \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'})$ with
  $\alpha_{s'} = \min D_x \wedge \delta_{s'} \not\in D_x$.
\end{definition}

\begin{lemma}[\label{def:fbfse}FB+FS+E $\centernot\implies$ DCP]
  Forwarding of best up-to-date, forwarding of staleness, and echos is
  not sufficient to guarantee dynamic consistent partitioning.
\end{lemma}

\begin{proof}
Stale control information as stated in Definition~\ref{def:lwo} may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. We simplify the echo since $\alpha_c^3$ is already transiting
towards $a$. Finally, Figure~\ref{fig:problemD} shows that eventually,
$c$ stays in the deleted partition $P_a$ for not having received
$\delta_a$ that $b$ blocked earlier. \end{proof}

As a cascading effect, in Figure~\ref{fig:problemD}, not only $c$
wrongfully believes that it belongs to $a$'s partition when $a$
already deleted it, but it could contribute to its inconsistency by
blocking farther but up-to-date messages from actual sources. To reach
consistent partitioning, \processes need
\begin{inparaenum}[(i)]
\item to detect possibly blocking conditions 
\item to purge all stale messages from the system
\item enabling up-to-date notifications to eventually reach every
  \process.
\end{inparaenum}

\begin{lemma}[\label{def:detection}Detector existence]
  Assuming a chain of \process $[x,\ldots, z]$ that delivered a
  message $\alpha_x$, if $x$ deletes its partition, either all
  \processes eventually deliver $\delta_s$, or a \process along the
  chain detects the blocking of the delete notification
\end{lemma}

\TODO{But cannot just delete. It must be deleted only downstream, see
  figure.}

@@

\begin{lemma}[FB $\wedge$ FD $\implies$ Detection] \TODO{meow!}
\end{lemma}

\begin{definition}[\TODO{Assuming purge, echo => DCP}]
\end{definition}
